Stability concerns in random matrices

### 2018-04-28 random_matrix_stability.Rmd
Add a bunch of stuff, and more notes

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
A_comp <- NULL;
A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
A_mat  <- matrix(data = A_dat, nrow = 1000);
C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
A_mat  <- A_mat * C_mat;
#gamma1 <- c(4,   rep(1 - 1/333,  999));
#gamma2 <- c(1/4, rep(1 + 1/1332, 999));
gamma1 <- c(18,   rep(1 - 17/999,  999));
gamma2 <- c(1/18, rep(1 + (17/18)/999, 999));
#gamma1 <- c(40,   rep(1 - 13/333,  999));
#gamma2 <- c(1/40, rep(1 + (39/40)/999, 999));
mu_gam <- mean(gammas);
diag(A_mat) <- -1;
A1     <- gamma1 * A_mat;
A0     <- gamma2 * A_mat;
A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

@@

>>>>

```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
A_comp <- NULL;
A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
A_mat  <- matrix(data = A_dat, nrow = 1000);
C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
A_mat  <- A_mat * C_mat;
#gamma1 <- c(4,   rep(1 - 1/333,  999));
#gamma2 <- c(1/4, rep(1 + 1/1332, 999));
gamma1 <- c(18,   rep(1 - 17/999,  999));
gamma2 <- c(1/18, rep(1 + (17/18)/999, 999));
#gamma1 <- c(40,   rep(1 - 13/333,  999));
#gamma2 <- c(1/40, rep(1 + (39/40)/999, 999));
mu_gam <- mean(gammas);
diag(A_mat) <- -1;
A0     <- gamma1 * A_mat;
A1     <- gamma2 * A_mat;
A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

@@

Note that the sum of the real components of all eigenvalues needs to equal the sum of the diagonal of a matrix [@Allesina2015]. So we can break the two haves of the matrix represented by **b** above into the upper half where the sum equals `r sum(diag(A1[1:500, 1:500]))`, and the lower half where the sum equals `r sum(diag(A1[501:1000, 501:1000]))`. We can then partition the eigenvalues in each half, $\lambda_{upper}$ and $\lambda_{lower}$.

We can also note that the values of the off diagonals in the upper half of the matrix are multiplied by 1.95, so $\sigma_{upper} = \sigma(1.95) = 1.95$ and $\sigma_{lower} = \sigma(0.05) = 0.05$. This effectively creates two circles, one with a radius of $1.95\sqrt{SC}$; this is shown explicitly below with the red circles.

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(A0_vm);
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(A1_vm);
A1vec       <- A1vec[is.na(A1vec) == FALSE];
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
plot(A0_r, A0_i, xlim = c(-19.5, 16.5), ylim = c(-19.5,16.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
vl <- seq(from = 0, to = 2*pi, by = 0.001);
x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
text(x = -16.5, y = 14, labels = "a", cex = 2);
abline(v = 0, col = "blue");
points(x = x0, y = y0, type = "l", lwd = 3);
points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
plot(A1_r, A1_i, xlim = c(-19.5, 16.5), ylim = c(-19.5,16.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
     yaxt = "n");
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vecU      <- as.vector(A1_vm[1:500,]);
A1vecU      <- A1vecU[is.na(A1vecU) == FALSE];
A1vecL      <- as.vector(A1_vm[501:1000,]);
A1vecL      <- A1vecL[is.na(A1vecL) == FALSE];
x0 <- sqrt(500) * sd(A1vecU) * cos(vl) + mean(diag(A1[1:500,1:500]));
y0 <- sqrt(500) * sd(A1vecU) * sin(vl);
x1 <- sqrt(500) * sd(A1vecL) * cos(vl) + mean(diag(A1[501:1000,501:1000]));
y1 <- sqrt(500) * sd(A1vecL ) * sin(vl);
text(x = -16.5, y = 14, labels = "b", cex = 2);
abline(v = 0, col = "blue");
points(x = x1, y = y1, type = "l", col = "orange", lwd = 2)
points(x = x0, y = y0, type = "l", lwd = 2, lty = "solid", col = "orange");

A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(A1_vm);
A1vec       <- A1vec[is.na(A1vec) == FALSE];
x3 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
y3 <- sqrt(1000) * sd(A1vec) * sin(vl);
points(x = x3, y = y3, type = "l", col = "red", lwd = 3, lty = "dashed");

mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
```

Note that the large orange is pushed leftward (red dashed circles are identical), though I am not entirely sure why yet. It should be noted that the upper-left and lower-right squares of the matrix should be partitionable. The remaining squares have off-diagonals pulled from matching distributions, so we should really be able to think of the whole thing as a combination of the upper and lower rows, I think.

Also note how the smaller circle is pulled rightwards (centre = `r mean(x0)`) while the larger circle is pulled leftwards (centre = `r mean(x1)`), while the overall centre remains the same (red dashed circle, centre = `r mean(x3)`).
